---
title: "Hi! I’m Wendy Fung-Wu 👋"
---

I’m a graduate student at CUNY Baruch College majoring in **Business Analyst** concentration in **Data Analyst**. I’m passionate about blending _technology, culture, and leadership_, and I believe every project is an opportunity to bring people together to achieve something meaningful.

📌 Feel free to connect with me in [LinkedIn](https://www.linkedin.com/in/wendy-fungwu) or shoot me an [email](mailto:wendyfungwu@outlook.com)!  

[📄 View My Resume](https://drive.google.com/file/d/1uvrLePLWDyGHBPuVF5yNMyNIye_MKAfE/view?usp=sharing){.btn .btn-primary target="_blank"}

```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.75028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```

